---
title: "Bombay Stock Exchange Index Visualization using Plotly"
author: "Shashank Sane"
date: "June 11, 2017"
output: ioslides_presentation

fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Synopsis

This project visualizes Bombay Stock Exchange market variation from year 2001 to year 2017 using the S&P BSE 100 index data. More Information about the S&P BSE100 can be found at: http://www.bseindia.com/indices/DispIndex.aspx?iname=BSE100&index_Code=22&page=5B6DA9E7-CD6D-48D3-880A-05D1E9E1AA56

The data for BSE 100 can be downloaded from: http://www.bseindia.com/indices/IndexArchiveData.aspx

The visualization has been done using the Plotly package in R, furthermore two additional methods of achieving the same visualization has been shown in presentation.

## Data load and data wrangling 1.. {.smaller}

In this slide we will focus on creating a clean data frame for our visualization.

```{r chunk1, cache = F, echo = T, message = F, warning = F, tidy = F,eval=T}
# load the required libraries
library(plotly)
library(tidyr)
library(dplyr)
library(lubridate)

```

```{r chunk10, echo = TRUE,comment=""}
# load the dataset from downloaded csv file.
dfBSE <- read.csv("BSE100.csv",header = TRUE,stringsAsFactors = FALSE
                  ,row.names = NULL)

## once the data frame is loaded it can be seen that there is a mismatch 
## in column names and an extra column with null values has been loaded.
head(dfBSE,3)

# to correct this we will modify the dataframe in next slide

```

##  data wrangling 2 and plotting{.smaller}

```{r chunk15, echo = TRUE}
# extract column names
names <- colnames(dfBSE[,2:6])

# remove last column
dfBSE <- dfBSE[,1:5]

# re apply column names
colnames(dfBSE) <- names

# convert Date to datetime object using lubridate
dfBSE$Date <- dmy(dfBSE$Date)

# gather all columns except Date column into two columns index and value
# store this in new data frame dfBSE.Modified
dfBSE.Modified <- gather(dfBSE,index,value,-Date)
```
create plot using plotly

```{r chunk25, echo = TRUE,eval=FALSE}
plot_ly(data=dfBSE.Modified,x=~Date,y=~value, color = ~index,mode="line")%>%
  layout(title = "S&P BSE 100 Index (2001-2017)",
    xaxis = list(title = "Year",showgrid = T),yaxis = list( 
    title = "BSE 100 Index Value") 
  )

```

## {.smaller }

```{r chunk30,  cache = F, echo = F, message = F, warning = F, tidy = F,eval=T}
plot_ly(data=dfBSE.Modified,x=~Date,y=~value, color = ~index,mode="line")%>%
  layout(title = "S&P BSE 100 Index (2001-2017)",
    xaxis = list(title = "Year",showgrid = T),yaxis = list( 
    title = "BSE 100 Index Value") 
  )
```


We can see from the above plot that in general there is an increasing trend to the stock market index, this is commensurate with average 7% growth rate in Indian GDP. Also visible is a local minima around Oct-2008 period which coincides with the global economic recession period.

## Alternative implementation {.smaller .build}

alternatively the same plot can be drawn using plotly functional
aspect which allows one plotly object to be passed to another plotly
function.
```{r chunk40, cache = F, echo = T, eval=F,fig.cap="BSE 100 Index"}
add_lines(plot_ly(data=dfBSE.Modified,x=~Date,y=~value, color = ~index))%>%
  layout(title = "S&P BSE 100 Index (2001-2017)",
    xaxis = list(title = "Year",showgrid = T),yaxis = list( 
    title = "BSE 100 Index Value",titlefont="Courier New") 
  )
```
The same plot can also be generated from original data frame dfBSE as shown below, here we pass the plot_ly object to add_lines() for adding lines corresponding to "High" and "Low":

```{r chunk45, cache = F, echo = T, eval=F}

# have added only high and low value
plot_ly(data=dfBSE,x=~Date,color = I("black"))%>%
  add_lines(y = ~High, color = I("blue"))%>%
  add_lines(y = ~Low, color = I("red"))%>%
  layout(title = "S&P BSE 100 Index (2001-2017)",
    xaxis = list(title = "Year",showgrid = T),yaxis = list( 
    title = "BSE 100 Index Value") 
  )%>%
  

```

## {.smaller}
This completes demonstration of basic presentation using Rmarkdown and plotly.
Additional information about R markdown ioslides presentation can be found at: http://rmarkdown.rstudio.com/ioslides_presentation_format.html#overview

Addition information for plotly with R can be found at:
https://plot.ly/r/


### **Thank You!!.**